--
title: "Seldin Report: Ssec QQ Plots"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Calculate Ssec scores for Secreted and Non-Secreted proteins v. full datasets
  a. Plot Ssec scores in a QQ plot
2. Calculate Ssec scores for Tissue-Exclusive Secreted and Tissue-Exclusive Non-Secreted proteins v. full datasets
  b. Plot Ssec scores in a QQ plot

Following the procedure laid out in Seldin, we are comparing secreted proteins to non-secreted proteins from the origin tissue, without the designation of tissue-specific. This was compared to the full protein-coding dataset for the target tissue.

### Secreted v. Non-Secreted

```{r, echo = FALSE, message = FALSE}

library(tidyverse)
library(reshape2)
library(WGCNA)

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
sec_Ssec_brain = read.csv("Sec-Full-Ssec_brain.csv")
sec_Ssec_liver = read.csv("Sec-Full-Ssec_liver.csv")
nsec_Ssec_brain = read.csv("NSec-Full-Ssec_brain.csv")
nsec_Ssec_liver = read.csv("NSec-Full-Ssec_liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

sec_Ssec_brain = fix.read.csv(sec_Ssec_brain)
sec_Ssec_liver = fix.read.csv(sec_Ssec_liver)
nsec_Ssec_brain = fix.read.csv(nsec_Ssec_brain)
nsec_Ssec_liver = fix.read.csv(nsec_Ssec_liver)
```

Below are the QQ plots for the Ssec values comparing secreted to non-secreted. Ideally, Ssec values would have higher values in secreted proteins, showing that the Ssec measure was quantifying communication between tissues. However, the graphs show the opposite, with non-secreted proteins having higher correlations to the target tissue proteins than the secreted proteins.  

```{r, echo = FALSE}

## BRAIN

sec = sec_Ssec_brain$Ssec
nsec = nsec_Ssec_brain$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Brain Proteins"))
abline(a = 0, b = 1)

## LIVER

sec = sec_Ssec_liver$Ssec
nsec = nsec_Ssec_liver$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Liver Proteins"))
abline(a = 0, b = 1)

```

We can confirm by looking at the maximum values for each designation.

```{r, echo = FALSE}

top_sec_brain = head(sec_Ssec_brain)
top_nsec_brain = head(nsec_Ssec_brain)
top_sec_liver = head(sec_Ssec_liver)
top_nsec_liver = head(nsec_Ssec_liver)

out = rbind(top_sec_brain$Ssec, top_nsec_brain$Ssec, top_sec_liver$Ssec, top_nsec_liver$Ssec)
rownames(out) = c("Secreted Brain", "Non-Secreted Brain", "Secreted Liver", "Non-Secreted Liver")

knitr::kable(out, caption = "Top Ssec Scores by Category")

```

### Tissue-Exclusive Secreted v. Tissue-Exclusive Non-Secreted

Below is the same procedure, but with only RNA sequences that are exclusive to the tissue of interest. Liver appears to favor secreted proteins a bit more, albeit with a much lower sample size. Brain, however, still favors non-secreted protein associations. We can try to repeat this procedure, changing target genes to also be tissue-exclusive, instead of all protein-coding genes.  

```{r, echo = FALSE}

`%!in%` <- Negate(`%in%`)

te.sec_Ssec_brain = sec_Ssec_brain[sec_Ssec_brain$Gene_symbol %!in% sec_Ssec_liver$Gene_symbol,]
te.sec_Ssec_liver = sec_Ssec_liver[sec_Ssec_liver$Gene_symbol %!in% sec_Ssec_brain$Gene_symbol,]
te.nsec_Ssec_brain = nsec_Ssec_brain[nsec_Ssec_brain$Gene_symbol %!in% nsec_Ssec_liver$Gene_symbol,]
te.nsec_Ssec_liver = nsec_Ssec_liver[nsec_Ssec_liver$Gene_symbol %!in% nsec_Ssec_brain$Gene_symbol,]


## BRAIN

sec = te.sec_Ssec_brain$Ssec
nsec = te.nsec_Ssec_brain$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Brain Proteins"))
abline(a = 0, b = 1)

## LIVER

sec = te.sec_Ssec_liver$Ssec
nsec = te.nsec_Ssec_liver$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Liver Proteins"))
abline(a = 0, b = 1)

```

```{r, echo = FALSE, eval = FALSE}

# server code for sec TE v. nsec TE

#### Secreted ####
## BRAIN ##
te.sec_brain = protein_brain[, colnames(protein_brain) %in% rownames(te.sec_Ssec_brain)]
target_liver = protein_liver

brain.liv = bicorAndPvalue(te.sec_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(sec_Ssec_brain[,1])
bot_brain = tail(sec_Ssec_brain[,1])


## LIVER ##

target_brain = protein_brain
te.sec_liver = protein_liver[, colnames(protein_liver) %in% rownames(te.sec_Ssec_liver)]

liv.brain = bicorAndPvalue(te.sec_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(sec_Ssec_liver[,1])
bot_liver = tail(sec_Ssec_liver[,1])

#### Not Secreted ####

## BRAIN ##
te.nsec_brain = protein_brain[, colnames(protein_brain) %in% rownames(te.nsec_Ssec_brain)]

brain.liv.c = bicorAndPvalue(te.nsec_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv.c$p))

nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
#top_brain = head(tsns_Ssec_brain[,1])
#bot_brain = tail(ts_Ssec_brain[,1])


## LIVER ##

te.nsec_liver = protein_liver[, colnames(protein_liver) %in% rownames(te.nsec_Ssec_liver)]

liv.brain.c = bicorAndPvalue(te.nsec_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain.c$p))

nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))



```

```{r, echo = FALSE, eval = FALSE}
# double checking
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
te.sec_Ssec_brain = read.csv("TE.Sec_Ssec_brain.csv")
te.sec_Ssec_liver = read.csv("TE.Sec_Ssec_liver.csv")
te.nsec_Ssec_brain = read.csv("TE.NSec_Ssec_brain.csv")
te.nsec_Ssec_liver = read.csv("TE.NSec_Ssec_liver.csv")

te.sec_Ssec_brain = fix.read.csv(te.sec_Ssec_brain)
te.sec_Ssec_liver = fix.read.csv(te.sec_Ssec_liver)
te.nsec_Ssec_brain = fix.read.csv(te.nsec_Ssec_brain)
te.nsec_Ssec_liver = fix.read.csv(te.nsec_Ssec_liver)

## BRAIN

sec = te.sec_Ssec_brain$Ssec
nsec = te.nsec_Ssec_brain$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Brain Proteins"))
abline(a = 0, b = 1)

## LIVER

sec = te.sec_Ssec_liver$Ssec
nsec = te.nsec_Ssec_liver$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Liver Proteins"))
abline(a = 0, b = 1)


```

```{r}
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
associate_brain = read.csv("Associations_Brain.csv")
associate_liver = read.csv("Associations_Liver.csv")

top_brain = head(te.sec_Ssec_brain[,1])
bot_brain = tail(te.sec_Ssec_brain[,1])

assoc_top_brain = associate_brain[associate_brain$Var1 %in% top_brain,]
top_assoc_top_brain = assoc_top_brain[assoc_top_brain$value < 0.05,]
one = top_assoc_top_brain[top_assoc_top_brain$Var1 == top_brain[1],]

top_liver = head(te.sec_Ssec_liver[,1])
bot_liver = tail(te.sec_Ssec_liver[,1])

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
write.csv(one, "OneSecretedProtein_LessThan5Per.csv")

```

### checking target shit

```{r, echo = FALSE}

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

ts_Ssec_target_brain = read.csv("Target_TESec_Ssec_brain.csv")
ts_Ssec_target_liver = read.csv("Target_TESec_Ssec_liver.csv")
tsns_Ssec_target_brain = read.csv("Target_TENSec_Ssec_brain.csv")
tsns_Ssec_target_liver = read.csv("Target_TENSec_Ssec_liver.csv")
## BRAIN

sec = ts_Ssec_target_brain$Ssec
nsec = tsns_Ssec_target_brain$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Brain Proteins"))
abline(a = 0, b = 1)

## LIVER

sec = ts_Ssec_target_liver$Ssec
nsec = tsns_Ssec_target_liver$Ssec

cdf = ecdf(sec)

quant = c()
for (i in sec)
{
  gee = cdf(i)
  quant = c(quant, gee)
}

vals.nsec = quantile(nsec, probs = quant)

this = cbind(c(rep("Secreted", times = length(sec)), rep("Not Secreted", times = length(vals.nsec))), c(sec, vals.nsec))
colnames(this) = c("trt", "val")
this = as.data.frame(this)
this$val = as.numeric(this$val)

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Liver Proteins"))
abline(a = 0, b = 1)

```

