---
title: "Seldin Report: Ssec Scores"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Report on Ssec scores from "Seldin Methods"
  a. Find distribution of Ssec values
  b. Find distribution of correlation p-values
  
* Using exclusive ts proteins unless we decide later to change back to Uniprot TS
  
```{r, echo = FALSE, message = FALSE}

library(tidyverse)
library(reshape2)
library(WGCNA)
#library(reshape)
#library(ggplot2)

# Data Input

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_liver = read.csv("Liver_WGCNA.csv")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")
sec_annot = read.csv("Sec_Annotation.csv")
tissuespec_protein_brain = read.csv("TissueSpecific_Protein_Brain.csv")
tissuespec_protein_liver = read.csv("TissueSpecific_Protein_Liver.csv")
ts_protein_brain = read.csv("TissueSpecific2_Protein_Brain.csv")
ts_protein_liver = read.csv("TissueSpecific2_Protein_Liver.csv")

fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_liver = fix.read.csv(protein_liver)
protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)
sec_annot = fix.read.csv(sec_annot)
tissuespec_protein_brain = fix.read.csv(tissuespec_protein_brain)
tissuespec_protein_liver = fix.read.csv(tissuespec_protein_liver)
ts_protein_brain = fix.read.csv(ts_protein_brain)
ts_protein_liver = fix.read.csv(ts_protein_liver)

```
Calculating Ssec values
```{r, echo = FALSE, message = FALSE}

#### BRAIN ####

brain.liv = bicorAndPvalue(secreted_protein_brain, protein_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

## tissue-specific genes

ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(ts_Ssec_brain[,1])
bot_brain = tail(ts_Ssec_brain[,1])


#### LIVER ####

liv.brain = bicorAndPvalue(secreted_protein_liver, protein_brain, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(liv.brain$p))

Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

## tissue-specific genes

ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(ts_Ssec_liver[,1])
bot_liver = tail(ts_Ssec_liver[,1])

```

PROCEED WITH CAUTION
```{r, eval = FALSE}
####################################
# Calc Ssec for all proteins# Don't do this !!

brain.liv = bicorAndPvalue(protein_brain, protein_liver, use = 'pairwise.complete.obs')
# brain.liv$p

scores = rowSums(-log(brain.liv$p))

Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

## tissue-specific genes

top_brain = head(Ssec_brain[,1])
```

```{r}
# Visualizations
#hist(Ssec_brain$Ssec)
#hist(ts_Ssec_brain$Ssec)

#Brain
h1 = cbind(rep("brain", times = length(Ssec_brain$Ssec)), Ssec_brain$Ssec)
h2 = cbind(rep("TS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

compare = rbind(h1, h2)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)


ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Secreted Proteins v. TS Proteins", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("liver", times = length(Ssec_liver$Ssec)), Ssec_liver$Ssec)
h2 = cbind(rep("TS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

compare = rbind(h1, h2)
colnames(compare) = c("trt", "val")
compare = as.data.frame(compare)
compare$trt = as.factor(compare$trt)
compare$val = as.numeric(compare$val)


ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Secreted Proteins v. TS Proteins", x = "Ssec", y = "Density")

```

The above graphs show the density distribution of Ssec scores of the secreted proteins and the tissue-specific proteins (a subset of secreted proteins). The shape of the distributions are not very different, indicating that tissue-specificity does not make it more likely that there will be a significant relationship with target-tissue proteins. Consider the relationship between the non-tissue-specific proteins and the target tissue expression. 

```{r}
##
# brain
plotting = melt(brain.liv$p, id.vars = "refseq")
limit = plotting[as.character(plotting$Var1) %in% c(top_brain, bot_brain),]

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="P-Value Density Curve: Top6 and Bottom6 Brain Proteins", x = "P-Value", y = "Density")

# liver
plotting = melt(liv.brain$p, id.vars = "refseq")
limit = plotting[as.character(plotting$Var1) %in% c(top_liver, bot_liver),]

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="P-Value Density Curve: Top6 and Bottom6 Liver Proteins", x = "P-Value", y = "Density")


```

The above two plots are the density distribution of p-values within the proteins with the highest and lowest Ssec scores. We can see that the top Ssec score proteins have a peak on the lower end of p-values, which makes sense given they have the highest Ssec scores. Are there a high number of low but not significantly low p-values that could inflate these? Are we discounting some relationships because there aren't a lot of connections, but a small number of strong connections? The log transformation of the pvalues should weigh the lowest p-values much higher. Don't want to entirely discount pvalues just under a certain threshold. 

Secreted v TS- don't change shape as you go from secreted to tissue specific. Does that mean tissue specific not more likely to have a relationship with target tissue? See if correlation with proteins across tissues is more significant than the same correlation with the same proteins within the tissue.... Would secreted proteins have an effect on its own tissue? Perhaps some kind of relationship
Pval. Do we care at all about the magnitude of the correlation? I know we're not interested in direction. 

```{r}
# brain
magn = brain.liv$bicor

plotting = melt(magn, id.vars = "refseq")
limit = rbind(plotting[as.character(plotting$Var1) %in% top_brain,], plotting[as.character(plotting$Var1) %in% bot_brain,])

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="Magnitude Density Curve: Top6 and Bottom6 Brain Proteins", x = "Correlation", y = "Density")

# liver
magn = liv.brain$bicor
plotting = melt(magn, id.vars = "refseq")
limit = rbind(plotting[as.character(plotting$Var1) %in% top_liver,], plotting[as.character(plotting$Var1) %in% bot_liver,])

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="Magnitude Density Curve: Top6 and Bottom6 Liver Proteins", x = "Correlation", y = "Density")

```

Using the same top and bottom proteins, the above graphs show the correlation from the bicor calculation. The curves that are wider at the bottom are the top proteins, indicating that the more significant proteins also have greater correlation coefficients. 

NEXT: permute null distribution of Ssec; Calculate Ssec with cutoff; maybe figure out how to run Ssec on non-secreted proteins; null distribution of p-values?

We permuted the target tissue strains and recalculated Ssec to
construct the null distribution and maintain the gene-gene correlation structures in each tissue. These tests typically required
1,000–1,000,000 permutations, listed as p values in Table S1.
The permutation tests were designed to specifically to allow generation of a statistic while maintaining correlation structure to
limit the influence of inflation of cross-tissue gene correlations
(discussed below). We note that in these tests the p values
used reflect a placeholder for the correlation coefficient, which
does not require independence. 


```{r, eval = FALSE}

#permute strain names in target tissue array
# bicor> pval> Ssec val
# repeat 1000- compare observed value to mean of permuted values
target_protein = protein_liver
perm_test = matrix(NA, ncol = 1000, nrow = 174)
perm_test = as.data.frame(perm_test)
rownames(perm_test) = ts_Ssec_brain[,1]

for (i in 1:1000)
{
  
perm = sample(rownames(target_protein), size = length(rownames(target_protein)), replace = FALSE)
target_protein = target_protein[perm,]

brain.liv = bicorAndPvalue(ts_protein_brain, target_protein, use = 'pairwise.complete.obs')

scores = rowSums(-log(brain.liv$p))

Ssec_brain_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

perm_test[,i] = Ssec_brain_perm[,2]
} 

## i dont get how this works...

pee = perm_test[rownames(perm_test) == top_brain[2],]
hist(as.numeric(pee))
Ssec_brain[Ssec_brain$Gene_symbol == top_brain[2],]
# ok all of the true values are inside the null distribution. Did i do something wrong??

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
write.csv(perm_test, "Permutation_Seldin_Report.csv")

```

