---
title: "Tissue-Exclusive Proteins Report"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Redo the flow chart for tissue exclusive proteins
2. Do a full report of tissue exclusive proteins

```{r, echo = FALSE, message = FALSE}
# Packages
library(tidyverse)

# Data Input
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
protein_brain = read.csv("Brain_WGCNA.csv")
secreted_protein_brain = read.csv("Secreted_Brain.csv")

protein_liver = read.csv("Liver_WGCNA.csv")
secreted_protein_liver = read.csv("Secreted_Liver.csv")

fin.r.squared_brain = read.csv("R_Squared_Brain.csv")
fin.r.squared_liver = read.csv("R_Squared_Liver.csv")

notsec_brain = read.csv("NotSecreted_Brain.csv")
notsec_liver = read.csv("NotSecreted_Liver.csv")


fix.read.csv = function(x)
{
  names = x[,1]
  gee = x[,2:dim(x)[2]]
  rownames(gee) = names
  return(gee)
}

protein_brain = fix.read.csv(protein_brain)
secreted_protein_brain = fix.read.csv(secreted_protein_brain)

protein_liver = fix.read.csv(protein_liver)
secreted_protein_liver = fix.read.csv(secreted_protein_liver)

fin.r.squared_brain = fix.read.csv(fin.r.squared_brain)
fin.r.squared_liver = fix.read.csv(fin.r.squared_liver)

notsec_liver = fix.read.csv(notsec_liver)
notsec_brain = fix.read.csv(notsec_brain)



```

```{r}
# Tissue exclusive, secreted proteins
`%!in%` <- Negate(`%in%`)

te_names_brain = colnames(secreted_protein_brain) %!in% colnames(secreted_protein_liver)
te.sec_protein_brain = secreted_protein_brain[, te_names_brain]

te_names_liver = colnames(secreted_protein_liver) %!in% colnames(secreted_protein_brain)
te.sec_protein_liver = secreted_protein_liver[, te_names_liver]

# tissue exclusive, nonsecreted proteins

te_names_brain = colnames(notsec_brain) %!in% colnames(notsec_liver)
te.nsec_protein_brain = notsec_brain[, te_names_brain]

te_names_liver = colnames(notsec_liver) %!in% colnames(notsec_brain)
te.nsec_protein_liver = notsec_liver[, te_names_liver]


```

```{r, echo = FALSE, message = FALSE}
# OPTION 1

# flow chart
# https://towardsdatascience.com/r-visualizations-flow-charts-in-r-4cfa7f783872
library(DiagrammeR)

grViz(diagram = "digraph flowchart { 
      node [fontname = Arial, shape = oval, color = Lavender, style = filled, cex = 40]
      tab1 [label = '@@1', fontsize = 100]
      tab2 [label = '@@2', fontsize = 100]
      tab3 [label = '@@3', fontsize = 100]
      tab4 [label = '@@4', fontsize = 100]
      tab5 [label = '@@5', fontsize = 100]
      tab6 [label = '@@6', fontsize = 100]
      tab7 [label = '@@7', fontsize = 100]
      tab8 [label = '@@8', fontsize = 100]
      
      edge[minlen = 4, penwidth = 8, arrowsize = 4]
      tab1 -> tab2;
      tab2 -> tab3;
      tab3 -> tab4;
      tab4 -> tab5;
      tab5 -> tab6;
      tab4 -> tab7;
      tab7 -> tab8}
      
      [1]: 'Full Cleaned Data: B(132 individuals, 18,385 genes), L(132 individuals, 15,365 genes)'
      [2]: 'Means of Strains: B(45 strains, 18,385 genes), L(45 strains, 15,365 genes)'
      [3]: 'RI Strains Only: B(30, 18,385), L(30, 15,365)'
      [4]: 'Protein-Coding Genes Only: B(30, 16,457), L(30, 14,109)'
      [5]: 'Secreted Protein-Coding Genes: B(30, 657), L(30, 818)'
      [6]: 'Tissue-Exclusive Secreted Protein-Coding Genes: B(30, 174), L(30, 35)'
      [7]: 'Non-Secreted Protein-Coding Genes: B(30, 15,800), L(30, 13,591)'
      [8]: 'Tissue-Exclusive Non-Secreted Protein-Coding Genes: B(30, 2,589), L(30, 371)' 
      ", width = 1000, height = 400)

```

Labels are of the format Tissue dataset (number of strains, number of genes), unless otherwise specified. B indicates the brain dataset, while L indicates the liver dataset. The full cleaned data came in the form number of RNA reads per individual rat. The strain means of the reads were calculated by averaging the RNA reads per gene for all rats within a strain. The resulting dataset had data for 45 strains. Classically inbred strains were removed from the dataset, leaving only recombinant inbred strains, for which we have phenotype data. Then, all non-protein coding-genes were removed, as identified by _.   MORE

```{r}

# OPTION 2

# flow chart
# https://towardsdatascience.com/r-visualizations-flow-charts-in-r-4cfa7f783872
library(DiagrammeR)

grViz(diagram = "digraph flowchart { 
      node [fontname = Arial, shape = oval, color = Lavender, style = filled]
      tab1 [label = '@@1', fontsize = 100]
      tab2 [label = '@@2', fontsize = 100]
      tab3 [label = '@@3', fontsize = 100]
      tab4 [label = '@@4', fontsize = 100]
      tab5 [label = '@@5', fontsize = 100]
      tab6 [label = '@@6', fontsize = 100]

      
      edge[minlen = 4, penwidth = 8, arrowsize = 4]
      tab1 -> tab2;
      tab2 -> tab3;
      tab3 -> tab4;
      tab2 -> tab5;
      tab5 -> tab6;
}
      
      [1]: 'All RNA Reads: B(18,385 genes), L(15,365 genes)'
      [2]: 'Protein-Coding Genes Only: B(16,457), L(14,109)'
      [3]: 'Secreted Protein-Coding Genes: B(657), L(818)'
      [4]: 'Tissue-Exclusive Secreted Protein-Coding Genes: B(174), L(35)'
      [5]: 'Non-Secreted Protein-Coding Genes: B(15,800), L(13,591)'
      [6]: 'Tissue-Exclusive Non-Secreted Protein-Coding Genes: B(2,589), L(371)' 
      ", width = 1000, height = 400)




```

```{r}

library(Gmisc)
library(glue)
library(htmlTable)
library(grid)
library(magrittr)
```

```{r}

# https://cran.r-project.org/web/packages/Gmisc/vignettes/Grid-based_flowcharts.html

grid.newpage()
txt1 <-
"Full Cleaned Data
Brain (132 ind, 18,385 genes)
Liver (132 ind, 15,365 genes)"
boxFC = boxGrob(txt1, x = 0.15, y = .89)
txt2 = 
  "Means of Strains
Brain (45 strains, 18,385 genes)
Liver (45 strains, 15,365 genes)"
boxMS = boxGrob(txt2, x = 0.15, y = 0.65)
txt3 = 
  "RI Strains Only
Brain (30 strains, 18,385 genes)
Liver (30 strains, 15,365 genes)"
boxRI = boxGrob(txt3, x = 0.15, y = 0.40)
txt4 = 
  "Protein-Coding Genes Only
Brain (30 strains, 16,457 genes)
Liver (30 strains, 14,109 genes)"
boxPC = boxGrob(txt4, x = 0.15, y = 0.15)
txt5 = 
  "Secreted Protein-Coding Genes
Brain (30 strains, 657 genes)
Liver (30 strains, 818 genes)"
boxSe = boxGrob(txt5, x = 0.5, y = 0.89)
txt6 = 
  "Tissue-Exclusive Secreted 
  Protein-Coding Genes
Brain (30 strains, 174 genes)
Liver (30 strains, 35 genes)"
boxTES = boxGrob(txt6, x = 0.5, y = 0.65)
txt7 = 
  "Non-Secreted Protein-Coding Genes
Brain (30 strains, 15,800 genes)
Liver (30 strains, 13,591 genes)"
boxNSe = boxGrob(txt7, x = 0.5, y = 0.40)
txt8 = 
  "Tissue-Exclusive Non-Secreted 
  Protein-Coding Genes
Brain (30 strains, 2,589 genes)
Liver (30 strains, 371 genes)"
boxTENS = boxGrob(txt8, x = 0.5, y = 0.13)

connectGrob(boxFC, boxMS, "vertical")
connectGrob(boxMS, boxRI, "vertical")
connectGrob(boxRI, boxPC, "vertical")
connectGrob(boxPC, boxSe, "Z", "r")
connectGrob(boxSe, boxTES, "vertical")
connectGrob(boxPC, boxNSe, "Z", "r")
connectGrob(boxNSe, boxTENS, "vertical")

boxFC
boxMS
boxRI
boxPC
boxSe
boxTES
boxNSe
boxTENS




```

```{r}

grid.newpage()
txt1 =
"All Normalized Expression Values
Brain (18,385 genes)
Liver (15,365 genes)"
boxAR = boxGrob(txt1, x = 0.15, y = 0.5)
txt4 = 
  "Protein-Coding Genes Only
Brain (16,457 genes)
Liver (14,109 genes)"
boxPC = boxGrob(txt4, x = 0.35, y = 0.5)
txt5 = 
  "Secreted Protein-Coding Genes
Brain (657 genes)
Liver (518 genes)"
boxSe = boxGrob(txt5, x = 0.35, y = 0.70)
txt6 = 
  "Tissue-Exclusive, Secreted 
  Protein-Coding Genes
Brain (174 genes)
Liver (35 genes)"
boxTES = boxGrob(txt6, x = 0.55, y = 0.70)
txt7 = 
  "Non-Secreted Protein-Coding Genes
Brain (15,800 genes)
Liver (13,591 genes)"
boxNSe = boxGrob(txt7, x = 0.35, y = 0.30)
txt8 = 
  "Tissue-Exclusive, Non-Secreted 
  Protein-Coding Genes
Brain (2,577 genes)
Liver (368 genes)"
boxTENS = boxGrob(txt8, x = 0.55, y = 0.30)
txt9 = 
  "Tissue-Exclusive 
  Protein-Coding Genes
Brain (2,751 genes)
Liver (403 genes)"
boxTE = boxGrob(txt8, x = 0.55, y = 0.5)

connectGrob(boxAR, boxPC, "horizontal")
connectGrob(boxPC, boxSe, "vertical")
connectGrob(boxSe, boxTES, "horizontal")
connectGrob(boxPC, boxNSe, "vertical")
connectGrob(boxNSe, boxTENS, "horizontal")
connectGrob(boxPC, boxTE, "horizontal")

boxAR
boxPC
boxSe
boxTES
boxNSe
boxTENS
boxTE




```